\section{The \JoinCalc}

\subsubsection{Syntax}

We use \joincalc as described by Fournet et al.
\cite{fournet_calculus_1996}.
Assuming an infinite set of channel names
$ \mathcal{N} = \{ x, y, z, u, v, \kappa, \ldots \} $,
we use
$ \tilde{x}, \tilde{y}, \ldots $
for tuples of names
and define the syntax as follows.

% left-align explanations?
\begin{align}
  \tag*{\textbf{processes}}
  \\
  P, Q
  \grmr \jsnd{x}{\tilde{v}}   \tag*{emission of $\tilde{v}$ on $x$}
  \altn \nonumber \jdef{D}{P} \tag*{definition of D in P}
  \altn P \jpar Q             \tag*{parallel composition}
  \altn \jnullproc            \tag*{empty process}
  \\ \nonumber
  \\
  \tag*{\textbf{definitions}}
  \\
  D, E
  \grmr \jrct{J}{P} \tag*{elementary clause}
  \altn D \jcon E   \tag*{simultaneous definitions}
  \altn \jnulldef   \tag*{empty definition}
  \\ \nonumber
  \\
  \tag*{\textbf{join-patterns}}
  \\
  J, K
  \grmr \jrcv{x}{\tilde{v}} \tag*{reception of $\tilde{v}$ on x}
  \altn J \jpat K           \tag*{composed join-pattern}
\end{align}

The simplest possible process is the empty process $\jnullproc$.
Also, we can send a tuple of names $\tilde{v}$ on a channel $x$
and compose any two processes $P$ and $Q$ in parallel.
The most interesting construct is the definition $\jdef{D}{P}$,
which persistently defines a number of elementary clauses.
Each elementary clause $\jrct{J}{R}$ consists of
a join-pattern $J$,
composed from receptions
$\jrcv{x_1}{\tilde{v}_1} \jpat \cdots \jpat \jrcv{x_n}{\tilde{v}_n}$
and a process R.
When a message is available on each channel $x_1, \ldots, x_n$,
they can be consumed atomically,
which bindes them to the variables $\tilde{v}_1, \ldots, \tilde{v}_n$
and starts the process R.
THe defined names $x_1, \ldots, x_n$ are available in both $D$ and $P$
(so they can be used recursively).
Note that all received names in a join-pattern must be distinct,
making both patterns
$ \jrcv{x}{u,u} $ and
$ \jrcv{x}{u} \jpat \jrcv{y}{u} $
invalid.

We will now define the scopes of received, defined and free variables
($\oprv$, $\opdv$, $\opfv$).

\begin{alignat*}{2}
  & \rv{\jrcv{x}{\tilde{v}}} & &= \{ u \in \tilde{v} \} \\
  & \rv{J \jpat K}           & &= \rv{J} \uplus \rv{K} \\
  \\
  & \dv{\jrcv{x}{\tilde{v}}} & &= \{ x \} \\
  & \dv{J \jpat K}           & &= \dv{J} \cup \dv{K} \\
  \\
  & \dv{\jrct{J}{P}}         & &= \dv{J} \\
  & \dv{D \jcon E}           & &= \dv{D} \cup \dv{E} \\
  & \dv{\jnulldef}           & &= \emptyset \\
  \\
  & \fv{\jrct{J}{P}}         & &= \dv{J} \cup (\fv{P} - \rv{J}) \\
  & \fv{D \jcon E}           & &= \fv{D} \cup \fv{E} \\
  & \fv{\jnulldef}           & &= \emptyset \\
  \\
  & \fv{\jsnd{x}{v}}         & &= \{ x \} \cup \{ u \in \tilde{v} \} \\
  & \fv{\jdef{D}{P}}         & &= (\fv{P} \cup \fv{D}) - \dv{D} \\
  & \fv{P \jpat Q}           & &= \fv{P} \cup \fv{Q} \\
  & \fv{\jnullproc}          & &= \emptyset \\
\end{alignat*}


\subsubsection{Semantics}

The semantics of \joincalc is described using the reflexive chemical abstract
machine (RCHAM) \cite{fournet_reflexive_1996}.
A RCHAM configuration
$ \rchamcnf{\mathcal{R}}{\mathcal{M}} $
consists of a multiset of active definitions $\mathcal{R}$,
called ``reactions",
and a multiset of active processes $\mathcal{M}$,
called ``molecules".

There are multiple reversible heating/cooling rules $\rchamheating$,
corresponding to the structural congruence between processes,
and a single reduction rule $\rchamreduction$.
In each of these rules, there is an implicit unchanged context of reactions and
molecules which we omit for brevity.

\begin{align*}
  \rchamequ{str-join}{
    \rchamcnf*{}{P \jpar Q}
  }{
    \rchamcnf*{}{P, Q}
  }
  \\
  \rchamequ{str-null}{
    \rchamcnf*{}{\jnullproc}
  }{
    \rchamcnf*{}{}
  }
  \\
  \rchamequ{str-and}{
    \rchamcnf*{D \jcon E}{}
  }{
    \rchamcnf*{D, E}{}
  }
  \\
  \rchamequ{str-nodef}{
    \rchamcnf*{\jnulldef}{}
  }{
    \rchamcnf*{}{}
  }
  \\
  \rchamequ{str-def}{
    \rchamcnf*{}{\jdef{D}{P}}
  }{
    \rchamcnf*{D\sigma_{\opdv}}{P\sigma_{\opdv}}
  }
  \\
  \\
  \rchamred{red}{
    \rchamcnf*{\jrct{J}{P}}{J\sigma_{\oprv}}
  }{
    \rchamcnf*{\jrct{J}{P}}{P\sigma_{\oprv}}
  }
\end{align*}

In (str-dev), the substitution $\sigma_{\opdv}$ instantiates the defined
variables $\dv{D}$ to distinct, fresh names that are not free in any reaction
or molecule in the configuration.
In (red), $\sigma_{\oprv}$ simply substitutes the received variables $\rv{J}$.
Note that (red) makes use of the syntactic similarity between join-patterns
and parallel emission of messages in processes.

To apply a reduction, we usually first need to dissolve processes
using the heating rules until we can apply (red).
Afterwards, the solution can be cooled down again
(see example \ref{example_cham}).

We can obtain a reduction relation $\joinreduction$
on processes (instead of RCHAM configurations)
by defining for any two processes P and P':
\begin{equation*}
  P \joinreduction P'
  \ \ \textbf{iff} \ \ 
  \rchamcnf{\emptyset}{P}
  \rchamheating^*\rchamreduction\rchamheating^*
  \rchamcnf{\emptyset}{P'}
\end{equation*}

\begin{example}
\label{example_cham}
We can follow the reduction of a simple process in the RCHAM.
\begin{align*}
  &&\rchamcnfset*
    {}
    {\jdef
      {\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}
      {\jsnd{x}{a} \jpar \jsnd{x}{b} \jpar \jsnd{y}{c}}}
  \\
  \rchamheating\ \tag{str-def}
  &&\rchamcnfset*
    {\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}
    {\jsnd{x}{a} \jpar \jsnd{x}{b} \jpar \jsnd{y}{c}}
  \\
  \rchamheating\ \tag{str-join}
  &&\rchamcnfset*
    {\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}
    {\jsnd{x}{a}, \jsnd{x}{b} \jpar \jsnd{y}{c}}
  \\
  \rchamreduction \tag{red}
  &&\rchamcnfset*
    {\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}
    {\jsnd{x}{a}, P\substitution{\subst{b}{u}, \subst{c}{v}}}
  \\
  \rchamheating\ \tag{str-join}
  &&\rchamcnfset*
    {\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}
    {\jsnd{x}{a} \jpar P\substitution{\subst{b}{u}, \subst{c}{v}}}
  \\
  \rchamheating\ \tag{str-def}
  &&\rchamcnfset*
    {}
    {\jdef
      {\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}
      {\jsnd{x}{a} \jpar P\substitution{\subst{b}{u}, \subst{c}{v}}}}
\end{align*}

and thus

\begin{equation*}
  \jdef
    {\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}
    {\jsnd{x}{a} \jpar \jsnd{x}{b} \jpar \jsnd{y}{c}}
  \ \joinreduction\ 
  \jdef
    {\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}
    {\jsnd{x}{a} \jpar P\substitution{\subst{b}{u}, \subst{c}{v}}}
\end{equation*}

Note the non-determinism in receiving $b$ instead of $a$ on $x$.
Instead of
$ \jsnd{x}{a} \jpar P\substitution{\subst{b}{u}, \subst{c}{v}} $
we could also arrive at
$ \jsnd{x}{b} \jpar P\substitution{\subst{a}{u}, \subst{c}{v}} $.
\end{example}

\begin{example}\ \\
\TODO{3cm}{example (lock)?}
\end{example}


\subsection{The \CoreJoinCalc}

The full calculus can be encoded in a fragment of itself, called core calculus
\cite{fournet_reflexive_1996}.
Since all of its processes are also valid processes in the full calculus,
it suffices to describe its semantics using the RCHAM as shown above,
only using the rules (str-join), (str-def) and (red).

\begin{align*}
  P, Q
  \grmr \jsnd{x}{u}
  \altn P \jpar Q
  \altn \jdef{\jrct{\jrcv{x}{u} \jpat \jrcv{y}{v}}{P}}{Q}
\end{align*}


\subsection{The \DistJoinCalc}

In addition to an infinite set of channel names
$ \mathcal{N} = \{ x, y, z, \kappa, \ldots \} $,
we now use an infinite set of location names
$ \mathcal{L} = \{ a, b, c, \ldots \} $.
These can also be sent and received, which means that values
$ u, v, \ldots $ in $ \jsnd{x}{\tilde{v}} $
can now be either channel or location names.
We extend the syntax of \joincalc by two new constructs
\cite{fournet_calculus_1996}.
Locations can be introduced using a new kind of definition:

\begin{align*}
  D, E
  \grmr \dots
  \altn \jloc{a}{D}{P}
\end{align*}

This initializes a new sublocation $a$ with definitions $D$
and a process $P$ residing there.
We demand two additional syntactic restrictions:
Firstly, location $a$ can only be defined once in any definition $D$.
Secondly, channel $x$ may only appear in the join-patterns of one location.
These conditions make the following definitions invalid:

\begin{align*}
  &\jdef
    {\jloc{a}{D}{P} \jcon \jrct{\jloc{a}{E}{Q}}{R}}
    {S}
  \\
  &\jdef
    {\jloc{a}{\jrct{\jrcv{x}{u}}{P}}{Q}
     \jcon
     \jloc{b}{\jrct{\jrcv{x}{v}}{R}}{S}
     }
    {T}
\end{align*}

Also, we provide a construct for migration:

\begin{align*}
  P, Q
  \grmr \dots
  \altn \jgo{b}{\kappa}
\end{align*}

It migrates the current location to $b$ (becoming a sublocation)
and afterwards emits an empty message on $\kappa$.

The semantics of the new constructs are defined using a distributed version
of the RCHAM, called DRCHAM.
It is a multiset of solutions
$ \drchamcnf{\varphi}{\mathcal{R}}{\mathcal{M}} $
with location \emph{labels}
$ \varphi, \psi, \ldots \in \mathcal{L}^* $,
separated by $||$.
We consider trees of locations.
For each location $a$ in the tree,
the DRCHAM contains a solution with a label
consisting (left to right) of the path from the root down to $a$.
This means that $\varphi$ is a sublocation of $\psi$
if $\psi$ is a prefix of $\varphi$.

Each single solution internally evolves by the rules introduced before
(with the addition that (str-def) also produces fresh location names).
Furthermore, we add three new rules for the two new syntactical constructs:

\begin{align*}
  \drchamequ{str-loc}{(a \ \text{frozen)}}{
    \drchamcnf{\varphi}{\jloc{a}{D}{P}}{}
  }{
    \drchamcnf{\varphi}{}{}
    \drchampar
    \drchamcnf{\varphi a}{D}{P}
  }
  \\
  \drchamred{comm}{(x \in \dv{J})}{
    \drchamcnf{\varphi}{}{\jsnd{x}{\tilde{v}}}
    \drchampar
    \drchamcnf{}{\jrct{J}{P}}{}
  }{
    \drchamcnf{\varphi}{}{}
    \drchampar
    \drchamcnf{}{\jrct{J}{P}}{\jsnd{x}{\tilde{v}}}
  }
  \\
  \drchamred{move}{}{
    \drchamcnf{\varphi}{\jloc{a}{D}{P \jpar \jgo{b}{\kappa}}}{}
    \drchampar
    \drchamcnf{\psi b}{}{}
  }{
    \drchamcnf{\varphi}{}{}
    \drchampar
    \drchamcnf{\psi b}{\jloc{a}{D}{P \jpar \jsnd{\kappa}{}}}{}
  }
\end{align*}

In heating direction, (str-loc) initializes a new machine.
The side condition ``a frozen" applies in the cooling direction
and expresses that there may be no solutions at sublocations of $a$
(with a label containing $a$).
This enforces that the rule is applied to all sublocations of a first
to freeze the whole subtree, preparing migration.
(comm) allows a message on a channel that is located on another location
to move to that location, where it can be received using (red) later.
The last rule (move) gives the semantics of migration: $a$ becomes a sublocation
of $b$. The emitted $\jsnd{\kappa}{}$ can be used to delay execution of a
process until migration is completed.
